---
title: "User Guide"
author: "Antoine Champion"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{magrittr}
---

```{r include = FALSE}
library(optional)
```

## Introduction ##

This package adds an `optional` type, similar to `Option` in F#, OCaml and Scala, to `Maybe` in Haskell, and to nullable types in `C#`.

It should be used instead of `NULL` for values that might be missing or otherwise invalid.

This package also introduces pattern matching.

## Using the optional type ##

`optional` is an object wrapper which indicates whether the object is valid or not.

### Declaring an optional object ###

An optional variable can be set to `some(object)` or to `none`.

```{r}
a <- some(5)
class(a)
```

Operators and print will have the same behavior with an `optional` than with its base type.
```{r}
a == 5
a
```

Note that `some(some(obj))` equals `some(obj)` and that `some(none)` equals `FALSE`.

### Optionals on functions ###

Given a function `f()`, to handle properly `optional` arguments and wraps its return  type into an `optional`, one should use `make_opt()` the following way:

```
f_opt <- make_opt(f)
```

1. Every `optional` argument passed to `f_opt()` will be converted to its original type before being sent to `f()`. If one or more of them is `none`, several behaviors are available (see `?make_opt`).
2. If `f()` returns null, or if an error is thrown during its execution, then `f_opt()` returns `none`. Else it will return `some(f(...))`.

For instance:

```{r}
c_opt <- make_opt(c)
c_opt(some(2), none, some(5))
c_opt()
```
	
## Pattern matching ##

Patterns are used in many functional languages in order to process variables in an exhaustive way.

The syntax is the following:

```
match_with( variable,
pattern , result-function,
...
```

If `variable` matches a `pattern`, `result-function` is called.
For comparing optional types, it is a better habit to use `match_with()` rather than a conditional statement.

```{r}
library(magrittr)

a <- 5
match_with(a,
  . %>% some(.),          print,
  none,                   function() print("Error!")
)
```

1. Each `pattern` can be either:

    * an object or a primitive type (direct comparison with `variable`)
    * a list (match if `variable` is in the list)
    * a `magrittr` functional sequence that matches if it returns `variable`. The dot `.` denotes the variable to be matched.

2. If `result-function` takes no arguments, it will be called as is. Else, the only argument that will be sent is `variable`.

```{r}
a <- 5
match_with(a,
  1,                      function() print("Matched exact value"),
  list(2, 3, 4),          function(x) paste("Matched in list:", x),
  . %>% if (. > 4).,      function(x) paste("Matched in condition:", x)
)
```